---
title: "Cost-Effectiveness Analysis: Time to Appropriate Antibiotic Therapy (TTAA)"
subtitle: "Assessing the Economic Value of Rapid Diagnostics in Bloodstream Infections"
author: "Research Team"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    number-sections: true
    code-fold: show
    theme: cosmo
---

# Introduction

This document contains the source code for a cost-effectiveness model assessing the value of reducing the Time to Appropriate Antibiotic Therapy (TTAA) for hospitalised patients with bloodstream infections (BSI). 

The analysis compares **Standard Care** (Delayed TTAA) vs. **Intervention** (Early TTAA, facilitated by rapid diagnostic technologies). The model uses a decision-analytic framework combining a short-term acute phase (30-day survival) with a long-term Markov extrapolation.

# Setup and Dependencies

We begin by loading the necessary libraries for matrix operations, data manipulation, parallel processing, and visualization.


```{r}
#| label: setup

## Clean memory and load packages
rm(list = ls())
# Matrix exponentials and mathematical functions
library(expm) 
# Data manipulation and functional programming
library(dplyr)
library(tidyr)
library(purrr)
# Visualization
library(ggplot2)
library(patchwork)
# Parallel processing for PSA
library(future)
library(furrr)
library(progressr)

# Set up multi-core parallel processing locally
# Reserves 2 cores for system stability, capping usage at a reasonable limit
n_cores <- min(parallel::detectCores() - 6, 14) 
cat("Using", n_cores, "cores for parallel processing\n")
plan(multisession, workers = n_cores)
handlers(global = TRUE)
handlers("progress")
```
# Parameters and transformations
## Hazard Ratio conversion
The clinical literature often reports the hazard ratio (HR) of mortality per hour of delay in antibiotic administration. However, specific diagnostic technologies reduce TTAA by discrete intervals (e.g., 2 hours, 4 hours).
The function below converts a per-hour HR into a "per-interval" HR (e.g., HR for a 4-hour reduction) and scales the standard error accordingly using the Delta method.
```{r}
#| label: 0.1 Transform per-hour HR to per-interval HR
getPerIntervalHR <- function(perHourHR,
                             perHourHRSE, 
                             DeltaHours = 1) {
  intHR   = perHourHR ^ DeltaHours
  intHRSE = DeltaHours * ( perHourHR ^ (DeltaHours - 1) ) * perHourHRSE

  out <- list(intHR = intHR, intHRSE = intHRSE)
  out
}
```

## Model parameters
This section defines the base case values for the model. These include clinical probabilities, utility weights, costs, and structural assumptions.
```{r}
#| label: 0.2 Global Parameter Definition (change assumptions as needed)
params <- list(
  # --- PSA Configuration ---
  N = 2000,                  # Number of Monte Carlo simulations

  # --- Clinical Parameters ---
  dTTAA_hours    = 4,        # Reduction in time to therapy provided by the technology
  
  # Baseline (Delayed Group) Mortality
  p30_del_mean   = 0.40,     # 30-day mortality probability
  p30_del_se     = 0.015,
  
  # Effectiveness
  perHourHR      = 0.92,     # HR of mortality per hour reduction in TTAA
  perHourHRSE    = 0.04,     # Standard Error of the HR
  
  # Note: HR_mean and HR_se are placeholders on a per hour basis; they are calculated dynamically below based on dTTAA_hours
  HR_mean        = 0.89,     
  HR_se          = 0.04,

  # Long-term Annual Mortality (Conditional on 30-day survival)
  pLT_del_mean = 0.15,       # Delayed group
  pLT_del_se   = 0.03,
  pLT_ear_mean = 0.15,       # Early group (Assumed same as delayed in base case)
  pLT_ear_se   = 0.03,

  # --- Health State Utilities (QALYs) ---
  u_acute_mean = 0.60,       # Utility during the acute infection phase
  u_acute_se   = 0.01,
  u_LT_mean    = 0.80,       # Utility during the long-term post-acute phase
  u_LT_se      = 0.02,

  # --- Structural Parameters ---
  horizon_y          = 10,   # Time horizon in years
  dt_acute_days      = 1,    # Cycle length for acute phase (days)
  dt_LT_years        = 1,    # Cycle length for long-term phase (years)
  discRate           = 0.03, # Discount rate for costs and outcomes

  # --- Cost Parameters (CHF) ---
  # Hospitalization
  cost_perHospDay_mean    = 975,
  costMultiplier_HospDay  = 0.30, # Used to define variance in PSA (Gamma dist)
  
  # Length of Stay (LOS) Impact
  baseline_LOS_days  = 15,      # Baseline LOS for delayed group survivors
  dLOS_perHour_mean  = -0.134,  # Reduction in LOS days per hour of TTAA reduction
  dLOS_perHour_se    = 0.008,
  
  # Technology Costs
  tech_cost_total         = 500000, # Fixed investment cost
  n_patients_using_tech   = 5000,   # Volume of patients over horizon
  tech_cost_per_patient   = NA,     # Calculated below
  
  # Long-term Management Costs (Annual)
  LT_cost_delayed_mean    = 0,      
  LT_cost_delayed_se      = 1000,
  LT_cost_early_mean      = 0,
  LT_cost_early_se        = 1000,
  costMultiplier_LT       = 0.30,

  # WTP threshold
  WTP                     = 100000
  )

# Derived Parameters: Convert per-hour HR to the specific interval HR
hrk <- getPerIntervalHR(params$perHourHR, params$perHourHRSE, params$dTTAA_hours)
params$HR_mean <- hrk$intHR
params$HR_se   <- hrk$intHRSE

# Derived Parameters: Calculate per-patient technology cost
if (is.na(params$tech_cost_per_patient)) {
  params$tech_cost_per_patient <- params$tech_cost_total / params$n_patients_using_tech
}

```

# Model structure

## Hazard and transition probability functions
These functions generate the Transition Probability Matrices (TPM) used in the Markov traces
```{r}
#| label: 1.1 Hazard Rate Conversion
# Use 30-day mortality to obtain a per-day hazard rate, under assumption of constant hazard over 30 days
hazard_p30delayed <- function(p30_delayed, days = 30) {
    if (p30_delayed <= 0 | p30_delayed >= 1) stop("p30_delayed is a probability and must be in (0,1).")
        -log(1-p30_delayed) /days
}
```

```{r}
#| label: 1.2 Transition Probability Matrices
# Create TP matrices for acute phase (e.g. here 30 days) and long-term phase
# s: probablity of surivival in the next time step dt, given survival up to time t; i.e. no event between t and {t+dt} given survival up to t.
acuteTPM <- function(h, dt_days = 1) {
    s <- exp(-h * dt_days)
    matrix(c(s, 1-s, 
             0, 1),
           nrow = 2, byrow = T,
           dimnames = list(c("alive", "dead"), c("alive", "dead"))
           )
           }

LT_TPM <- function(p_yearly, dt_years = 1) { 
    if (p_yearly <0 || p_yearly >= 1) stop("p_yearly is a probability and must be in [0,1).")
    s_yearly <- 1 - p_yearly
    s_dt <- s_yearly^dt_years # per-cycle survival (dt_years = 1/12 for monthly; = 1 for yearly)
    matrix(c(s_dt, 1-s_dt, 
             0, 1),
           nrow = 2, byrow = T,
           dimnames = list(c("alive", "dead"), c("alive", "dead"))
           )
           }

```

# Health Outcomes Calculation
## Acute Phase (First 30 Days)
This function calculates survival curves, expected life-days, and QALYs accrued during the initial hospitalization period.

```{r}
#| label: 1.3 Calculate expected time alive and QALYs in acute phase
haz30 <- hazard_p30delayed(0.3)
TPM30 <- acuteTPM(haz30)

acute_outcomes <- function(h, T_days = 30, dt_acute_days = 1, u_acute = 0.7) {
    n_cycles <- T_days / dt_acute_days
    P <- acuteTPM(h, dt_acute_days)
    I2 <- diag(2)
    dimnames(I2)[2] <- dimnames(P)[2]

    # Build a list of P to powers: P^0==I, P^1, P^2, P^3, ..., P^N, so total length of list is N+1
    P_n <- accumulate(
        .x = seq_len(n_cycles),
        .init = I2,
        .f = ~ .x %*% P
        )

    # Initial distribution of states: alives and deads, row vector i.e. 1, 0: all alive
    p_states0 <-matrix(c(1,0), nrow = 1, dimnames = list(NULL, c("alive", "dead")))

    # Start of cycle distributions dead/alive NB: P_n has length 31. dist_states0 * P^k for k = 0,...,N-1 (so index 1 is P^0 and index 31 is P^30)
    p_states_beginning <- map(P_n[1:n_cycles], ~ p_states0 %*% .x)

    # End of horizon distributions: dist_states0 * P^N
    p_states_T <- p_states0 %*% P_n[[n_cycles+1]]

    # Extract vector of proportion of alives; and end states
    p_alive_beginning <- map_dbl(p_states_beginning, ~ .x[,"alive"])
    p_alive_T <- p_states_T[,"alive"]
    p_dead_T <- p_states_T[,"dead"]

    # Expected time alive in acute phase; QALYs accumulated over acute phase
    exp_days_alive <- map_dbl(p_states_beginning, ~ .x[1,1]) |> sum() *dt_acute_days
    exp_years_alive <- exp_days_alive /365
    qaly_acute <- u_acute * exp_years_alive
    
    list(
    p_alive_beginning = p_alive_beginning,
    p_alive_T = p_alive_T,
    p_dead_T = p_dead_T,
    exp_days_alive_acute = exp_days_alive,
    exp_years_alive_acute = exp_years_alive,
    qaly_acute = qaly_acute
    )
    }
```

## Long-Term Phase (Post-30 Days)
This function models the trajectory of survivors from the acute phase over the remaining time horizon.

```{r}
#| label: 1.4 Calculate post-acute long-term time life-years and QALYs
# h_longTerm_annual is the hazard rate (assumed constant) that applies to survivors of the BSI post the acute phase
# Need to fill in the u_LT with something reasonable instead of 1

longTerm_outcomes <- function(p_longTerm_annual, horizon_y = 10, dt_LT_years = 1, discRate = 0.03, u_LT = 1) {
    n_cycles <- ceiling(horizon_y / dt_LT_years)
    P <- LT_TPM(p_longTerm_annual, dt_LT_years)
    I2 <- diag(2)
    dimnames(I2)[2] <- dimnames(P)[2]

    P_n <- accumulate(
    .x = seq_len(n_cycles),
    .init = I2,
    .f = ~ .x %*% P
    )

    # Beginning of cycles distributions
    p_states0 <-matrix(c(1,0), nrow = 1, dimnames = list(NULL, c("alive", "dead")))
    p_states_beginning <- map(P_n[1:n_cycles], ~ p_states0 %*% .x)
    p_alive_beginning <- map_dbl(p_states_beginning, ~ .x[,"alive"])

    # Discounting life time
    discFactor_percycle <- 1 / ((1 + discRate)^dt_LT_years) # per-cycle discount factor
    v_discFact <- discFactor_percycle^(0 : (n_cycles - 1)) # vector of DFs for each time step, starting at year 1 when no discouting (power 0)
    disc_timeAlivePost <- p_alive_beginning * v_discFact
    sum_disc_timeAlivePost <- sum(disc_timeAlivePost)
    LY_LT <- sum_disc_timeAlivePost * dt_LT_years # * by dt_LT_years is converting untis from months (unit of dt) to years!
    qaly_LT <- LY_LT * u_LT

    list(
    p_alive_beginning = p_alive_beginning,
    exp_years_alive_LT = LY_LT,
    qaly_LT = qaly_LT,
    v_discFact = v_discFact
    )
}

```

# Cost Calculations
This function aggregates costs from all sources: acute hospitalization, technology implementation, and long-term care. It accounts for the differential Length of Stay (LOS) driven by early treatment.

```{r}
#| label: cost-calculations
calculate_costs <- function(
    # Acute phase parameters
    exp_days_alive_acute_delayed,
    exp_days_alive_acute_early,
    dLOS_perHour,
    dTTAA_hours,
    cost_perHospDay,
    
    # Technology costs
    tech_cost_per_patient,
    
    # Long-term parameters
    p_alive_T_delayed,
    p_alive_T_early,
    p_alive_beginning_LT_delayed,
    p_alive_beginning_LT_early,
    LT_cost_delayed_annual,
    LT_cost_early_annual,
    
    # Model structure
    dt_LT_years,
    discRate,
    horizon_y
    ) {
    
    # Calculate number of long-term cycles
    n_cycles_LT <- ceiling(horizon_y / dt_LT_years)
    
    # Discount factors for costs (same as for QALYs)
    discFactor_percycle <- 1 / ((1 + discRate)^dt_LT_years)
    v_discFact <- discFactor_percycle^(0 : (n_cycles_LT - 1))
    
    # ACUTE PHASE COSTS (Hospitalization)
    baseline_LOS_days <- 10  # Baseline length of stay
    
    # Actual LOS for each group
    LOS_delayed <- baseline_LOS_days
    LOS_early <- baseline_LOS_days + (dLOS_perHour * dTTAA_hours) # Should be negative reduction due to dLOS_perHour <= 0; pmax to avoid negative values if baseline LOS is low
    
    # Hospitalization costs
    hosp_cost_delayed <- LOS_delayed * cost_perHospDay
    hosp_cost_early <- LOS_early * cost_perHospDay
    
    # TECHNOLOGY COSTS
    # Only early group uses the new technology; default late tech is 0 but can input costs if known
    tech_cost_delayed <- 0
    tech_cost_early <- tech_cost_per_patient
    
    # LONG-TERM COSTS
    # Discounted long-term costs: first per survivor, then actual costs accounting for proporion that survived at end of acute phase
    if (length(p_alive_beginning_LT_delayed) > 0) {
        # Annual costs multiplied by probability alive and discount factor (scalars recycled to length equal n_cycles)
        LT_costs_delayed_disc <- sum(p_alive_beginning_LT_delayed * LT_cost_delayed_annual * v_discFact * dt_LT_years)
        LT_costs_early_disc <- sum(p_alive_beginning_LT_early * LT_cost_early_annual * v_discFact * dt_LT_years)
        
        # Adjust for probability of surviving acute phase
        LT_costs_delayed_total <- p_alive_T_delayed * LT_costs_delayed_disc
        LT_costs_early_total <- p_alive_T_early * LT_costs_early_disc
    } else {
        LT_costs_delayed_total <- 0
        LT_costs_early_total <- 0
    }
    
    # TOTAL COSTS BY ARM
    total_cost_delayed <- hosp_cost_delayed + tech_cost_delayed + LT_costs_delayed_total
    total_cost_early <- hosp_cost_early + tech_cost_early + LT_costs_early_total
    
    # COST DIFFERENCES (early - delayed)
    d_hosp_costs <- hosp_cost_early - hosp_cost_delayed
    d_tech_costs <- tech_cost_early - tech_cost_delayed
    d_LT_costs <- LT_costs_early_total - LT_costs_delayed_total
    d_total_costs <- total_cost_early - total_cost_delayed
    
    list(
        # Individual cost components for each arm
        hosp_cost_delayed = hosp_cost_delayed,
        tech_cost_delayed = tech_cost_delayed,
        LT_cost_delayed = LT_costs_delayed_total,
        total_cost_delayed = total_cost_delayed,
        
        hosp_cost_early = hosp_cost_early,
        tech_cost_early = tech_cost_early,
        LT_cost_early = LT_costs_early_total,
        total_cost_early = total_cost_early,
        
        # Differences (early - delayed)
        d_hosp_costs = d_hosp_costs,
        d_tech_costs = d_tech_costs,
        d_LT_costs = d_LT_costs,
        d_total_costs = d_total_costs,
        
        # Additional outputs
        LOS_delayed = LOS_delayed,
        LOS_early = LOS_early,
        d_LOS_days = LOS_early - LOS_delayed
    )
}
```

# Preparing arm-specific health outcomes prior to contrasting


```{r}
#| label: 2.1 Outcomes per arm: delayed (baseline) vs early therapy
outcomes_perArm <- function(p30_delayed, 
                            HR_earlyVSdelayed, 
                            pLT_del_annual = 0.15,
                            pLT_ear_annual = 0.15,
                            horizon_y = 10, 
                            dt_acute_days = 1,
                            dt_LT_years = 1,
                            discRate = 0.03,
                            u_acute = 0.7, 
                            u_LT = 1,
                            cost_perHospDay = 1250,
                            dLOS_perHour = -0.134,
                            dTTAA_hours = 1,
                            tech_cost_per_patient = 500,
                            LT_cost_delayed_annual = 5000,
                            LT_cost_early_annual = 5000) {
    
    # Arm-specific acute phase outcomes
    h_delayedTTAA <- hazard_p30delayed(p30_delayed, days = 30)
    h_earlyTTAA <- h_delayedTTAA * HR_earlyVSdelayed

    acuteOutcomesDelayed <- acute_outcomes(h_delayedTTAA, T_days = 30, dt_acute_days, u_acute)
    acuteOutcomesEarly <- acute_outcomes(h_earlyTTAA, T_days = 30, dt_acute_days, u_acute)

    # Arm-specific long-term outcomes
    LTdelayed <- longTerm_outcomes(pLT_del_annual, horizon_y, dt_LT_years, discRate, u_LT)
    LTearly <- longTerm_outcomes(pLT_ear_annual, horizon_y, dt_LT_years, discRate, u_LT)

    # Arm-specific QALYs
    qaly_delayed <- acuteOutcomesDelayed$qaly_acute + 
        u_LT * (acuteOutcomesDelayed$p_alive_T * LTdelayed$exp_years_alive_LT)
    qaly_early <- acuteOutcomesEarly$qaly_acute + 
        u_LT * (acuteOutcomesEarly$p_alive_T * LTearly$exp_years_alive_LT)

    # Calculate costs
    costs <- calculate_costs(
        exp_days_alive_acute_delayed = acuteOutcomesDelayed$exp_days_alive_acute,
        exp_days_alive_acute_early = acuteOutcomesEarly$exp_days_alive_acute,
        dLOS_perHour = dLOS_perHour,
        dTTAA_hours = dTTAA_hours,
        cost_perHospDay = cost_perHospDay,
        tech_cost_per_patient = tech_cost_per_patient,
        p_alive_T_delayed = acuteOutcomesDelayed$p_alive_T,
        p_alive_T_early = acuteOutcomesEarly$p_alive_T,
        p_alive_beginning_LT_delayed = LTdelayed$p_alive_beginning,
        p_alive_beginning_LT_early = LTearly$p_alive_beginning,
        LT_cost_delayed_annual = LT_cost_delayed_annual,
        LT_cost_early_annual = LT_cost_early_annual,
        dt_LT_years = dt_LT_years,
        discRate = discRate,
        horizon_y = horizon_y
    )

    list(HR_earlyVSdelayed = HR_earlyVSdelayed,
        h_delayedTTAA,
        h_earlyTTAA,
        S30_delayed = acuteOutcomesDelayed$p_alive_T,
        S30_early = acuteOutcomesEarly$p_alive_T,
        qaly_delayed = qaly_delayed,
        qaly_early = qaly_early,
        pLT_del_annual,
        pLT_ear_annual,
        LYdelayed = LTdelayed$exp_years_alive_LT,
        LYearly = LTearly$exp_years_alive_LT,
        costs = costs
    )
    }
```

# Calculate contrasts

```{r}
#| label: 2.2 Generate contrast: early vs delayed TTAA; generate NMB and ICER

contrastsTTAA <- function(p30_delayed, 
                        HR_earlyVSdelayed, 
                        pLT_del_annual = 0.15,
                        pLT_ear_annual = 0.15,
                        horizon_y = 10, 
                        dt_acute_days = 1,
                        dt_LT_years = 1,
                        discRate = 0.03,
                        u_acute = 0.7, 
                        u_LT = 1,
                        dLOS_perHour,
                        cost_perHospDay,
                        dTTAA_hours,
                        tech_cost_per_patient = 500,
                        LT_cost_delayed_annual = 5000,
                        LT_cost_early_annual = 5000,                        
                        WTP = NULL) {
                            
    outcomes2Arms <- outcomes_perArm(
      p30_delayed               = p30_delayed,
      HR_earlyVSdelayed         = HR_earlyVSdelayed,
      pLT_del_annual            = pLT_del_annual,
      pLT_ear_annual            = pLT_ear_annual,
      horizon_y                 = horizon_y,
      dt_acute_days             = dt_acute_days,
      dt_LT_years               = dt_LT_years,
      discRate                  = discRate,
      u_acute                   = u_acute,
      u_LT                      = u_LT,
      cost_perHospDay           = cost_perHospDay,
      dLOS_perHour              = dLOS_perHour,
      dTTAA_hours               = dTTAA_hours,
      tech_cost_per_patient     = tech_cost_per_patient,
      LT_cost_delayed_annual    = LT_cost_delayed_annual,
      LT_cost_early_annual      = LT_cost_early_annual      
    )

    d_qalys <- outcomes2Arms$qaly_early - outcomes2Arms$qaly_delayed
    d_S30 <- outcomes2Arms$S30_early - outcomes2Arms$S30_delayed
    costs <- outcomes2Arms$costs

    out <- tibble(
        p30_delayed, HR_earlyVSdelayed, u_acute, u_LT, pLT_del_annual, pLT_ear_annual, 
        horizon_y, dt_acute_days, dt_LT_years, discRate,
        
        # Effectiveness outcomes
        d_qalys = d_qalys,
        d_S30 = d_S30,
        qaly_delayed = outcomes2Arms$qaly_delayed,
        qaly_early = outcomes2Arms$qaly_early,
        S30_delayed = outcomes2Arms$S30_delayed,
        S30_early = outcomes2Arms$S30_early,
        
        # Time to treatment
        dTTAA_hours = dTTAA_hours,
        
        # Length of stay
        dLOS_perHour = dLOS_perHour,
        LOS_delayed = costs$LOS_delayed,
        LOS_early = costs$LOS_early,
        d_LOS_days = costs$d_LOS_days,
        
        # Cost components - by arm
        cost_perHospDay = cost_perHospDay,
        hosp_cost_delayed = costs$hosp_cost_delayed,
        tech_cost_delayed = costs$tech_cost_delayed,
        LT_cost_delayed = costs$LT_cost_delayed,
        total_cost_delayed = costs$total_cost_delayed,

        hosp_cost_early = costs$hosp_cost_early,
        tech_cost_early = costs$tech_cost_early,
        LT_cost_early = costs$LT_cost_early,
        total_cost_early = costs$total_cost_early,
        
        # Cost differences
        d_hosp_costs = costs$d_hosp_costs,
        d_tech_costs = costs$d_tech_costs,
        d_LT_costs = costs$d_LT_costs,
        d_total_costs = costs$d_total_costs
    )

    if (!is.null(WTP)) {
        out <- out |> mutate(
            WTP = WTP,
            headroom = WTP * d_qalys,
            NMB = WTP * d_qalys - d_total_costs,
            ICER = d_total_costs / d_qalys,
            CE_decision = if_else(ICER <= WTP, "Cost-effective", "Not cost-effective")
        )
    }
    out
}
```

# Probabilistic sensitivity analysis

## Assumptions on how to describe sampling uncertainty around key parameters
PSA allows us to account for parameter uncertainty. Instead of using point estimates, we draw parameters from probability distributions.Sampling DistributionsWe employ the Method of Moments to fit standard distributions to our input parameters (mean and standard error):Beta Distribution: Used for probabilities (bounded 0 to 1).Log-Normal Distribution: Used for Hazard Ratios (bounded > 0, skewed).Gamma Distribution: Used for costs (bounded > 0, skewed). For the Gamma distribution, we fit the shape ($k$) and scale ($\theta$) parameters such that specific quantiles (2.5% and 97.5%) align with a defined variance (e.g., mean $\pm$ 30%).
```{r}
#| label: 4.1 PSA Distribution Helper Functions
# Build PSA helper function for probabilistic draws; input sample mean and SE that will be treated as actual mean and SD of distribution
draw_Beta_MOM <- function(N, theta, SEtheta) {
    a <- theta * ( theta * (1 - theta) /SEtheta^2 - 1)
    b <- ( 1 - theta ) * ( theta * (1 - theta) / SEtheta^2 - 1)
    rbeta(n = N, shape1 = a, shape2 = b)
    }


drawHR_Normal_MOM <- function(N, meanHR, SEHR) {
    SD_logNormal = log((SEHR/meanHR)^2 +1) |> sqrt()
    mean_logNormal = log(meanHR) - 0.5 * (SD_logNormal)^2
    HR_draws <- rlnorm(N, meanlog = mean_logNormal, sdlog = SD_logNormal)
    HR_draws
    }


draw_Gamma_withMean <- function(N, mu_gamma, meanMultiplier) {
    quantiles <- c(0.025, 0.975)
    target_qvalues <- c(mu_gamma * (1 - meanMultiplier), mu_gamma * (1 + meanMultiplier))

    # insert log of k just to prevent negative values of k (when re-exponentiatied) since k must be positive
    getSSE <- function(lnk) {
            k = exp(lnk)
            theta = mu_gamma / k
            draw_qgamma <- qgamma(quantiles, shape = k, scale = theta)
            SSE <- sum((target_qvalues - draw_qgamma)^2)
            SSE
            }
    
    lnk_hat <- optimize(getSSE, interval = c(-20, 20))$minimum
    k <- exp(lnk_hat)
    theta <- mu_gamma / k
    # Now parameterise gamma dist and draw
    rgamma(N, shape = k, scale = theta)
    }

```

## PSA execution

This function generates the random draws for all parameters and executes the model in parallel for $N$ iterations.

```{r}
#| label: 4.2 PSA Execution Function
# PSA function that will run on contrastsTTAA function
psa_contrastsTTAA <- function(
    N,
    p30_del_mean, p30_del_se,
    HR_mean, HR_se,
    pLT_del_mean, pLT_del_se,
    pLT_ear_mean, pLT_ear_se,
    u_acute_mean, u_acute_se,
    u_LT_mean, u_LT_se,
    horizon_y, dt_acute_days, dt_LT_years, discRate,
    dTTAA_hours,
    dLOS_perHour_mean, dLOS_perHour_se,
    cost_perHospDay_mean, costMultiplier_HospDay,
    tech_cost_per_patient,
    LT_cost_delayed_mean, LT_cost_delayed_se,
    LT_cost_early_mean, LT_cost_early_se,
    costMultiplier_LT,
    WTP = NULL) {
         # Clinical parameters
        p30_delayed = draw_Beta_MOM(N, p30_del_mean, p30_del_se)
        HR_earlyVSdelayed = drawHR_Normal_MOM(N, meanHR = HR_mean, SEHR = HR_se)
        pLT_del_annual = draw_Beta_MOM(N, pLT_del_mean, pLT_del_se)
        pLT_ear_annual = draw_Beta_MOM(N, pLT_ear_mean, pLT_ear_se)

        # Length of stay impact
        dLOS_perHour = rnorm(N, dLOS_perHour_mean, dLOS_perHour_se)

        # Utilities
        u_acute = draw_Beta_MOM(N, u_acute_mean, u_acute_se)
        u_LT = draw_Beta_MOM(N, u_LT_mean, u_LT_se)

        # COSTS
        # Hospitalisation costs per day
        cost_perHospDay = draw_Gamma_withMean(N, mu_gamma = cost_perHospDay_mean, meanMultiplier = costMultiplier_HospDay)
        # Long-term costs: deterministic version
        # LT_cost_delayed_annual = rep(LT_cost_delayed_mean, N)
        # LT_cost_early_annual = rep(LT_cost_early_mean, N)

        # Long-term costs: stochastic version, using Gamma distribution
        LT_cost_delayed_annual = draw_Gamma_withMean(N, mu_gamma = LT_cost_delayed_mean, meanMultiplier = costMultiplier_LT)
        LT_cost_early_annual = draw_Gamma_withMean(N, mu_gamma = LT_cost_early_mean, meanMultiplier = costMultiplier_LT)

        # Technology costs: for now deterministic
        tech_cost_per_patient = rep(tech_cost_per_patient, N)

        
        draws <- tibble(
            p30_delayed               = p30_delayed,
            HR_earlyVSdelayed         = HR_earlyVSdelayed,
            pLT_del_annual            = pLT_del_annual,
            pLT_ear_annual            = pLT_ear_annual,
            horizon_y                 = horizon_y,
            dt_acute_days             = dt_acute_days,
            dt_LT_years               = dt_LT_years,
            discRate                  = discRate,
            u_acute                   = u_acute,
            u_LT                      = u_LT,
            WTP                       = WTP,
            dTTAA_hours               = dTTAA_hours,
            cost_perHospDay           = cost_perHospDay,
            dLOS_perHour              = dLOS_perHour,
            tech_cost_per_patient     = tech_cost_per_patient,
            LT_cost_delayed_annual    = LT_cost_delayed_annual,
            LT_cost_early_annual      = LT_cost_early_annual
        )
        
        #psa_runs <- pmap_dfr(draws, contrastsTTAA)
        psa_runs <- future_pmap_dfr(draws, contrastsTTAA, .progress = TRUE)
        psa_runs
        }

```


# Results
## Run PSA

```{r}
#| label: 5.1 Running model with input parameters in section 0.0
results_psa <- psa_contrastsTTAA(
  N                      = params$N,
  p30_del_mean           = params$p30_del_mean,
  p30_del_se             = params$p30_del_se,
  HR_mean                = params$HR_mean,
  HR_se                  = params$HR_se,
  pLT_del_mean           = params$pLT_del_mean,
  pLT_del_se             = params$pLT_del_se,
  pLT_ear_mean           = params$pLT_ear_mean,
  pLT_ear_se             = params$pLT_ear_se,
  u_acute_mean           = params$u_acute_mean,
  u_acute_se             = params$u_acute_se,
  u_LT_mean              = params$u_LT_mean,
  u_LT_se                = params$u_LT_se,
  horizon_y              = params$horizon_y,
  dt_acute_days          = params$dt_acute_days,
  dt_LT_years            = params$dt_LT_years,
  discRate               = params$discRate,
  dTTAA_hours            = params$dTTAA_hours,
  dLOS_perHour_mean      = params$dLOS_perHour_mean,
  dLOS_perHour_se        = params$dLOS_perHour_se,
  cost_perHospDay_mean   = params$cost_perHospDay_mean,
  costMultiplier_HospDay = params$costMultiplier_HospDay,
  tech_cost_per_patient  = params$tech_cost_per_patient,
  LT_cost_delayed_mean   = params$LT_cost_delayed_mean,
  LT_cost_delayed_se     = params$LT_cost_delayed_se,
  LT_cost_early_mean     = params$LT_cost_early_mean,
  LT_cost_early_se       = params$LT_cost_early_se,
  costMultiplier_LT      = params$costMultiplier_LT,
  WTP                    = params$WTP
)
```

## Visualisation plots

```{r}
#| label: 6.1 Distribution plots
# Incremental QALYs
p1 <- ggplot(results_psa, aes(x = d_qalys)) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = mean(results_psa$d_qalys), 
             linetype = "dashed", color = "red", size = 1) +
  labs(x = "Incremental QALYs", y = "Frequency",
       title = "Distribution of Incremental QALYs") +
  theme_minimal() +
    theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12)
  )

# Incremental costs
p2 <- ggplot(results_psa, aes(x = d_total_costs)) +
  geom_histogram(bins = 50, fill = "coral", alpha = 0.7) +
  geom_vline(xintercept = mean(results_psa$d_total_costs), 
             linetype = "dashed", color = "red", size = 1) +
  labs(x = "Incremental Costs (CHF)", y = "Frequency",
       title = "Distribution of Incremental Costs") +
  theme_minimal() +
    theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12)
  )

# Net Monetary Benefit
p3 <- ggplot(results_psa, aes(x = NMB)) +
  geom_histogram(bins = 50, fill = "lightgreen", alpha = 0.7) +
  geom_vline(xintercept = 0, linetype = "solid", color = "black", size = 1) +
  geom_vline(xintercept = mean(results_psa$NMB), 
             linetype = "dashed", color = "darkgreen", size = 1) +
  labs(x = "Net Monetary Benefit (CHF)", y = "Frequency",
       title = paste0("NMB Distribution (WTP = CHF", params$WTP, ")")) +
  theme_minimal() +
    theme(
    plot.title = element_text(size = 24, face = "bold"),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    axis.text = element_text(size = 12)
  )

(p1 + p2) / p3

ggsave("./plots/Distributions.png", width = 10, height = 7, dpi = 300, bg = "white")

```

## Cost-Effectiveness Plane
Scatter plot of incremental costs versus incremental QALYs. The red dashed line represents the Willingness-to-Pay (WTP) threshold. Points below this line are cost-effective.

```{r}
#| label: 6.2 Cost-effectiveness plane

ggplot(results_psa, aes(x = d_qalys, y = d_total_costs)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_abline(slope = results_psa$WTP, intercept = 0, linetype = 2, color = "red") +
  geom_abline(slope = (results_psa$WTP * 0.25), intercept = 0, linetype = 2, color = "blue") +
  geom_point(alpha = 0.25, size = 1.2) +
  labs(title = "Cost-Effectiveness Plane",
       x = "Difference in QALYs (Early − Delayed)",
       y = "Difference in Costs (CHF, Early − Delayed)") +
  coord_cartesian(ylim = c(-800, 400)) +
    theme_minimal() +
    theme(
        plot.title = element_text(size = 24, face = "bold"),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text = element_text(size = 12)
    )

ggsave("./plots/CEplane.png", width = 10, height = 7, dpi = 300, bg = "white")
```

## Cost-Effectiveness Acceptability Curve (CEAC)
The CEAC shows the probability that the intervention is cost-effective at varying thresholds of Willingness-to-Pay.

```{r}
#| label: 6.3 Cost-effectiveness acceptability curve

# Calculate CEAC
wtp_range <- seq(0, 200000, by = 10)

# For each level of WTP, show the proportion of draws (in results_psa) that result in a positive MNB
ceac_data <- map_dfr(wtp_range, function(wtp) {
  results_psa |>
    mutate(ce = (wtp * d_qalys - d_total_costs) > 0) |>
    summarise(WTP = wtp, prob_CE = mean(ce))
})

# Plot CEAC
ggplot(ceac_data, aes(x = WTP, y = prob_CE)) +
  geom_line(color = "steelblue", size = 1.5) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray50") +
  geom_vline(xintercept = params$WTP, linetype = "dashed", color = "red") +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(x = "Willingness-to-Pay Threshold (CHF/QALY)",
       y = "Probability Cost-Effective",
       title = "Cost-Effectiveness Acceptability Curve") +
  theme_minimal() +
    theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12)
  )

ggsave("./plots/CEAC.png", width = 10, height = 7, dpi = 300, bg = "white")
```

## One-Way Sensitivity Analysis (Tornado Plot)
This section calculates the impact of varying single parameters on the Net Monetary Benefit (NMB). We use the 2.5th and 97.5th percentiles from the PSA distributions as the "Low" and "High" values for the deterministic sensitivity analysis.

```{r}
#| label: 6.4 Tornado plot, one-way sensitivity analysis (owsa) for the NMB
# Quantile funtions - recycle code from earlier but make output = quantile values
beta_qt <- function(theta, SEtheta, p = c(0.025, 0.975)) {
    a <- theta * ( theta * (1 - theta) /SEtheta^2 - 1)
    b <- ( 1 - theta ) * ( theta * (1 - theta) / SEtheta^2 - 1)
    qbeta(p, a, b)
    }

lnorm_qt <- function(meanHR, SEHR, p = c(0.025, 0.975)) {
  SD_logNormal = log((SEHR/meanHR)^2 +1) |> sqrt()
  mean_logNormal = log(meanHR) - 0.5 * (SD_logNormal)^2
  qlnorm(p, mean_logNormal, SD_logNormal)
}

norm_qt <- function(mean, se, p = c(0.025, 0.975)) {
  qnorm(p, mean, se)
}

# Run onsa on per hour HR rather than derived interval-HR, which also used Delta method to derive SEs
perHourHR_low  <- lnorm_qt(params$perHourHR, params$perHourHRSE)[1]
perHourHR_high <- lnorm_qt(params$perHourHR, params$perHourHRSE)[2]


# Parameter means or base values
param_base <- list(
  p30_delayed = params$p30_del_mean,
  HR_earlyVSdelayed = params$HR_mean,
  pLT_del_annual = params$pLT_del_mean,
  pLT_ear_annual = params$pLT_ear_mean,
  u_acute = params$u_acute_mean,
  u_LT = params$u_LT_mean,
  cost_perHospDay = params$cost_perHospDay_mean,
  dLOS_perHour = params$dLOS_perHour_mean,
  tech_cost_per_patient = params$tech_cost_per_patient,
  horizon_y = params$horizon_y,
  dt_acute_days = params$dt_acute_days,
  dt_LT_years = params$dt_LT_years,
  discRate = params$discRate,
  dTTAA_hours = params$dTTAA_hours,
  LT_cost_delayed_annual = params$LT_cost_delayed_mean,
  LT_cost_early_annual = params$LT_cost_early_mean,
  WTP = params$WTP
  )

baseNMB <- do.call(contrastsTTAA, param_base)$NMB

# Save parameters to vary in one way sensitivity analysis; tribble to write out row-wise
owsa_params <- tribble(
  ~parameter, ~label, ~low, ~high,
  "p30_delayed", "30-day mortality (delayed)", beta_qt(params$p30_del_mean, params$p30_del_se)[1], beta_qt(params$p30_del_mean, params$p30_del_se)[2],
  #"HR_earlyVSdelayed", "Hazard ratio (per hour)", lnorm_qt(params$HR_mean, params$HR_se)[1], lnorm_qt(params$HR_mean, params$HR_se)[2],  # this version runs ONSA on the "derived" interval-HR
  "HR_earlyVSdelayed", "Hazard ratio (per hour)", perHourHR_low ^ params$dTTAA_hours, perHourHR_high ^ params$dTTAA_hours, # this version runs ONSA on the original HR
  "pLT_del_annual", "Long-term mortality (delayed)", beta_qt(params$pLT_del_mean, params$pLT_del_se)[1], beta_qt(params$pLT_del_mean, params$pLT_del_se)[2],
  "pLT_ear_annual", "Long-term mortality (early)", beta_qt(params$pLT_ear_mean, params$pLT_ear_se)[1], beta_qt(params$pLT_ear_mean, params$pLT_ear_se)[2],
  "u_acute", "Utility (acute phase)", beta_qt(params$u_LT_mean, params$u_LT_se)[1], beta_qt(params$u_LT_mean, params$u_LT_se)[2],
  "u_LT", "Utility (long-term)", beta_qt(params$u_LT_mean, params$u_LT_se)[1], beta_qt(params$u_LT_mean, params$u_LT_se)[2],
  "dLOS_perHour", "LOS reduction (per hour)", norm_qt(params$dLOS_perHour_mean, params$dLOS_perHour_se)[1], norm_qt(params$dLOS_perHour_mean, params$dLOS_perHour_se)[2],
  "cost_perHospDay", "Hospital cost (per day)", params$cost_perHospDay_mean * 0.7, params$cost_perHospDay_mean * 1.3,
  "tech_cost_per_patient", "Technology cost (per patient)", params$tech_cost_per_patient * 0.1, params$tech_cost_per_patient * 10
)

# Produce spread between low and high values of NMB, per parametere varied
owsa_results <- owsa_params |>
  rowwise() |>
  mutate(
    NMB_low = { fnArgs <- param_base; fnArgs[[parameter]] <- low; do.call(contrastsTTAA, fnArgs)$NMB },
    NMB_high = { fnArgs <- param_base; fnArgs[[parameter]] <- high; do.call(contrastsTTAA, fnArgs)$NMB },
    spread = abs(NMB_high - NMB_low)
  ) |>
    ungroup()

# Tornado plot
ggplot(owsa_results, aes(y = reorder(label, spread))) +
  geom_segment(aes(x = NMB_low, xend = NMB_high, yend = label), linewidth = 6, color = "steelblue") +
  geom_vline(xintercept = baseNMB, linetype = "dashed", color = "red") +
  labs(x = "Net Monetary Benefit (CHF)", y = NULL,
       title = "Tornado Plot: one-way sensitivity analysis",
       subtitle = paste0("Mean NMB = CHF ", round(baseNMB, 0))) +
  theme_minimal()

ggsave("./plots/Tornado.png", width = 10, height = 7, dpi = 300, bg = "white")
```



# Stratified Analysis and Equity
This section investigates how the cost-effectiveness of the intervention changes across different clinical subgroups, specifically comparing a hypothetical patient scenario: those with cancer vs. those without.

## Defining Clinical Scenarios
We model two hypothetical groups: cancer and non-cancer.

Cancer: Characterised by higher baseline mortality (immunocompromised status), longer baseline length of stay, and potentially lower effectiveness of the intervention.

```{r}
#| label: 7.1 Define group-specific parameters: Cancer vs Non-Cancer
# Key differences: Cancer patients have higher baseline mortality due to severity/malignancy 
# and longer baseline length of stay. 
# In addition, earlier TTAA may reduce LOS differentially.
# Long-term survival is also allowed to differ between the 2 groups.

# Cancer group parameters
params_cancer <- params
params_cancer$p30_del_mean <- 0.50      # higher baseline mortality than base case
params_cancer$perHourHR <- 0.97         # early TTAA less effective
params_cancer$baseline_LOS_days <- 18   # longer LOS
params_cancer$dLOS_perHour_mean <- params$dLOS_perHour_mean * 0.5 # reduction of LOS due to early TTAA, smaller than baseline
params_cancer$pLT_del_mean <- 0.07      # higher long-term mortality. Equiv to 50% over 10 years
params_cancer$pLT_ear_mean <- 0.07      # assumed unchanged

# Non-Cancer group parameters
params_noncancer <- params
params_noncancer$p30_del_mean <- 0.25      # lower baseline mortality than base case 
params_noncancer$perHourHR <- params$perHourHR  # unchanged
params_noncancer$baseline_LOS_days <- 12   # shorter LOS
params_noncancer$dLOS_perHour_mean <- params$dLOS_perHour_mean # reduction of LOS due to early TTAA, same as baseline
params_noncancer$pLT_del_mean <- 0.035     # lower long-term mortality. Equiv to 30% over 10 years
params_noncancer$pLT_ear_mean <- 0.035     # assumed unchanged

# Recalculate derived intervention-HR (per X-hours) for each group
hrk_cancer <- getPerIntervalHR(params_cancer$perHourHR, params_cancer$perHourHRSE, params_cancer$dTTAA_hours)
params_cancer$HR_mean <- hrk_cancer$intHR
params_cancer$HR_se <- hrk_cancer$intHRSE

hrk_noncancer <- getPerIntervalHR(params_noncancer$perHourHR, params_noncancer$perHourHRSE, params_noncancer$dTTAA_hours)
params_noncancer$HR_mean <- hrk_noncancer$intHR
params_noncancer$HR_se <- hrk_noncancer$intHRSE
```

## Stratified PSA
```{r}
#| label: 7.2 PSA, by group
#| 
results_psa_cancer <- psa_contrastsTTAA(
  N                      = params_cancer$N,
  p30_del_mean           = params_cancer$p30_del_mean,
  p30_del_se             = params_cancer$p30_del_se,
  HR_mean                = params_cancer$HR_mean,
  HR_se                  = params_cancer$HR_se,
  pLT_del_mean           = params_cancer$pLT_del_mean,
  pLT_del_se             = params_cancer$pLT_del_se,
  pLT_ear_mean           = params_cancer$pLT_ear_mean,
  pLT_ear_se             = params_cancer$pLT_ear_se,
  u_acute_mean           = params_cancer$u_acute_mean,
  u_acute_se             = params_cancer$u_acute_se,
  u_LT_mean              = params_cancer$u_LT_mean,
  u_LT_se                = params_cancer$u_LT_se,
  horizon_y              = params_cancer$horizon_y,
  dt_acute_days          = params_cancer$dt_acute_days,
  dt_LT_years            = params_cancer$dt_LT_years,
  discRate               = params_cancer$discRate,
  dTTAA_hours            = params_cancer$dTTAA_hours,
  dLOS_perHour_mean      = params_cancer$dLOS_perHour_mean,
  dLOS_perHour_se        = params_cancer$dLOS_perHour_se,
  cost_perHospDay_mean   = params_cancer$cost_perHospDay_mean,
  costMultiplier_HospDay = params_cancer$costMultiplier_HospDay,
  tech_cost_per_patient  = params_cancer$tech_cost_per_patient,
  LT_cost_delayed_mean   = params_cancer$LT_cost_delayed_mean,
  LT_cost_delayed_se     = params_cancer$LT_cost_delayed_se,
  LT_cost_early_mean     = params_cancer$LT_cost_early_mean,
  LT_cost_early_se       = params_cancer$LT_cost_early_se,
  costMultiplier_LT      = params_cancer$costMultiplier_LT,
  WTP                    = params_cancer$WTP
) %>% 
  mutate(Group = "Cancer")

results_psa_noncancer <- psa_contrastsTTAA(
  N                      = params_noncancer$N,
  p30_del_mean           = params_noncancer$p30_del_mean,
  p30_del_se             = params_noncancer$p30_del_se,
  HR_mean                = params_noncancer$HR_mean,
  HR_se                  = params_noncancer$HR_se,
  pLT_del_mean           = params_noncancer$pLT_del_mean,
  pLT_del_se             = params_noncancer$pLT_del_se,
  pLT_ear_mean           = params_noncancer$pLT_ear_mean,
  pLT_ear_se             = params_noncancer$pLT_ear_se,
  u_acute_mean           = params_noncancer$u_acute_mean,
  u_acute_se             = params_noncancer$u_acute_se,
  u_LT_mean              = params_noncancer$u_LT_mean,
  u_LT_se                = params_noncancer$u_LT_se,
  horizon_y              = params_noncancer$horizon_y,
  dt_acute_days          = params_noncancer$dt_acute_days,
  dt_LT_years            = params_noncancer$dt_LT_years,
  discRate               = params_noncancer$discRate,
  dTTAA_hours            = params_noncancer$dTTAA_hours,
  dLOS_perHour_mean      = params_noncancer$dLOS_perHour_mean,
  dLOS_perHour_se        = params_noncancer$dLOS_perHour_se,
  cost_perHospDay_mean   = params_noncancer$cost_perHospDay_mean,
  costMultiplier_HospDay = params_noncancer$costMultiplier_HospDay,
  tech_cost_per_patient  = params_noncancer$tech_cost_per_patient,
  LT_cost_delayed_mean   = params_noncancer$LT_cost_delayed_mean,
  LT_cost_delayed_se     = params_noncancer$LT_cost_delayed_se,
  LT_cost_early_mean     = params_noncancer$LT_cost_early_mean,
  LT_cost_early_se       = params_noncancer$LT_cost_early_se,
  costMultiplier_LT      = params_noncancer$costMultiplier_LT,
  WTP                    = params_noncancer$WTP
) %>% 
  mutate(Group = "Non-Cancer")

# Bind 2 tibbles
results_psa_subgroups <- bind_rows(results_psa_cancer, results_psa_noncancer)
```

## Stratified visualisations

```{r}
#| label: 7.3 Comparing distributions by Patient Group
# QALYS
p4 <-ggplot(results_psa_subgroups, aes(x = d_qalys, fill = Group, color = Group)) +
  geom_histogram(aes(y = after_stat(density)), alpha = 0.4, position = "identity", bins = 50) +
  geom_density(alpha = 0, size = 1.2) +
  scale_fill_manual(values = c("Cancer" = "#E74C3C", "Non-Cancer" = "#3498DB")) +
  scale_color_manual(values = c("Cancer" = "#C0392B", "Non-Cancer" = "#2874A6")) +
  labs(x = "Incremental QALYs", y = "Density",
       title = "Incremental QALYs Distribution by Group",
       subtitle = "Histograms with density curves overlaid",
       fill = "Patient Group", color = "Patient Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        legend.position = "top")

# COSTS
p5 <- ggplot(results_psa_subgroups, aes(x = d_total_costs, fill = Group, color = Group)) +
  geom_histogram(aes(y = after_stat(density)), alpha = 0.4, position = "identity", bins = 50) +
  geom_density(alpha = 0, size = 1.2) +
  scale_fill_manual(values = c("Cancer" = "#E74C3C", "Non-Cancer" = "#3498DB")) +
  scale_color_manual(values = c("Cancer" = "#C0392B", "Non-Cancer" = "#2874A6")) +
  labs(x = "Incremental Costs", y = "Density",
       title = "Incremental Cost Distribution by Group",
       subtitle = "Histograms with density curves overlaid",
       fill = "Patient Group", color = "Patient Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        legend.position = "top")

# NMB 
p6 <- ggplot(results_psa_subgroups, aes(x = NMB, fill = Group, color = Group)) +
  geom_histogram(aes(y = after_stat(density)), alpha = 0.3, position = "identity", bins = 50) +
  geom_density(alpha = 0, size = 1.2) +  # Just the line, no fill
  geom_vline(xintercept = 0, linetype = "dashed", color = "black", size = 1) +
  scale_fill_manual(values = c("Cancer" = "#E74C3C", "Non-Cancer" = "#3498DB")) +
  scale_color_manual(values = c("Cancer" = "#C0392B", "Non-Cancer" = "#2874A6")) +
  labs(x = "Net Monetary Benefit (CHF)", y = "Density",
       title = "NMB Distribution by Group",
       subtitle = "Histograms with density curves overlaid",
       fill = "Patient Group", color = "Patient Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        legend.position = "top")

(p4 + p5) / p6

ggsave("./plots/Distributions_stratified.png", width = 10, height = 7, dpi = 300, bg = "white")
```

```{r}
#| label: 7.4 Cost-effectiveness plane by Patient Group
ggplot(results_psa_subgroups, aes(x = d_qalys, y = d_total_costs, color = Group)) +
  geom_hline(yintercept = 0, color = "gray50") +
  geom_vline(xintercept = 0, color = "gray50") +
  geom_abline(slope = params$WTP, intercept = 0, linetype = 2, color = "black", size = 1) +
  geom_point(alpha = 0.2, size = 1.5) +
  scale_color_manual(values = c("Cancer" = "#E74C3C", "Non-Cancer" = "#3498DB")) +
  labs(title = "Cost-Effectiveness Plane by Group",
       subtitle = paste0("WTP threshold: CHF ", format(params$WTP, big.mark = ",")),
       x = "Incremental QALYs (Early − Delayed)",
       y = "Incremental Costs (CHF, Early − Delayed)",
       color = "Patient Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        legend.position = "top")

ggsave("./plots/CEplane_stratified.png", width = 10, height = 7, dpi = 300, bg = "white")
```

```{r}
#| label: 7.5 CEAC by Patient Group

wtp_range <- seq(0, 200000, by = 1000)

ceac_subgroups <- map_dfr(wtp_range, function(wtp) {
  results_psa_subgroups %>%
    group_by(Group) %>%
    summarise(
      WTP = wtp,
      prob_CE = mean((wtp * d_qalys - d_total_costs) > 0),
      .groups = "drop"
    )
})

ggplot(ceac_subgroups, aes(x = WTP, y = prob_CE, color = Group)) +
  geom_line(size = 1.5) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray50") +
  geom_vline(xintercept = params$WTP, linetype = "dashed", color = "black") +
  scale_color_manual(values = c("Cancer" = "#E74C3C", "Non-Cancer" = "#3498DB")) +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(x = "Willingness-to-Pay Threshold (CHF/QALY)",
       y = "Probability Cost-Effective",
       title = "Cost-Effectiveness Acceptability Curves by Patient Group",
       color = "Patient Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        legend.position = "top")

ggsave("./plots/CEAC_stratified.png", width = 10, height = 7, dpi = 300, bg = "white")
```